---
title: "EDA"
author: "Alexandra Lawrence"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(broom)
library(patchwork)
library(knitr)
library(GGally)
library(tseries)
library(Kendall)
library(ggplot2)
```


```{r load-data, message = FALSE}
pico <- read.table("data/allpico-dataexport.txt", header = TRUE, 
                        stringsAsFactors = FALSE )
```

```{r remove-NAs}
pico[pico =="NaN"] <- NA
```

```{r convert-time}
pico1 <- pico %>% 
  unite("Date", c("Month", "Day", "Year"), sep = "/", remove = FALSE) %>% 
  arrange(Days)
```

## Exploring pH

```{r factor-NamedStorm}
pico1$NamedStorm <- as.factor(pico1$NamedStorm)
levels(pico1$NamedStorm) <- c("No", "Yes")
```

```{r plotting-pH}
ggplot(data=pico1, aes(x=Days, y= pHMEAN, color = NamedStorm)) +
  geom_point() +
  labs(title = "pHMEAN Plotted against Days")
```

Lower pH values observed when there is a named storm -- is this because of the storm or just a coincidence because pH tends to lower in warmer seasons and storms happen more often in summer?

```{r pH-model}
pH_model <- lm(pHMEAN ~ Days, data = pico1)
tidy(pH_model)
```

```{r calculations-min-ph}
smallest_pHMEAN <- pico1 %>% 
  filter(!is.na(pHMEAN))%>% 
  select(pHMEAN,Date) %>% 
  arrange(pHMEAN) %>% 
  slice(1)
smallest_pHMEAN
```

The smallest pH value was 9/20/2018 -- About a week after Hurricane Florence hit NC

```{r plotting-various-pH}
p1 <- ggplot(data=pico1, aes(x=Days, y= pHMEAN)) +
  geom_point() +
  labs(title = "pHMEAN vs. Days") +
  scale_y_continuous(limits=c(7, 8.5))
p2 <- ggplot(data=pico1, aes(x=Days, y= pHprobeMEAN)) +
  geom_point() +
  labs(title = "pHprobeMean vs. Days")
p3 <- ggplot(data=pico1, aes(x=Days, y= pH25CMEAN)) +
  geom_point() +
  labs(title = "pH25CMEAN vs. Days") + 
  scale_y_continuous(limits=c(7, 8.5))
```

```{r display-pH-with-patchwork}
p1+p3
```
\newpage

## Exploring Oxygen

```{r oxygen-plot}
p5 <- ggplot(data=pico1, aes(x = Days, y = OxygenMEAN)) +
  geom_point() +
  labs(title = "OxygenMEAN vs. Days") +
  scale_y_continuous(limits=c(5, 11.5))

p6 <- ggplot(data=pico1, aes(x = Days, y = OxygenProSoloMEAN)) +
  geom_point() +
  labs(title = "OxygenProSoloMEAN vs. Days") +
  scale_y_continuous(limits=c(5, 11.5))

p5+p6
```

```{r oxygen-sat-plot}
p10 <- ggplot(data=pico1, aes(x = Days, y = OxygenSatProSoloMEAN)) +
  geom_point() +
  labs(title = "OxygenSatProSoloMEAN vs. Days")

p11 <- ggplot(data=pico1, aes(x = Days, y = OxygenSaturationMEAN, color = NamedStorm)) +
  geom_point() +
  labs(title = "OxygenSaturationMEAN vs. Days")

p10+p11
```

```{r oxygen_sat_oxygen_temp}
ggplot(pico1) +
  geom_line(aes(y = TemperatureMEAN, x = Days),  size = 0.25) +
  geom_line(aes(y = OxygenMEAN, x = Days), color = "#c13d75ff")  + 
  geom_line(aes(y = OxygenSaturationMEAN, x = Days), color = "blue") +
  labs(title = "Mean Oxygen and Oxygen Saturation Plotted Against Temp")
```

```{r calculations-min-max-oxsat}
small_large_oxygen_sat <- pico1 %>% 
  filter(!is.na(OxygenSaturationMEAN))%>% 
  select(OxygenSaturationMEAN,Date, OxygenMEAN, TemperatureMEAN) %>% 
  arrange(OxygenSaturationMEAN) %>% 
  slice(1, 649)
small_large_oxygen_sat
```
Max mean oxygen saturation:  129.20% on 1/12/2020 -- cooler temp than min
Min mean oxygen saturation: 75.95% on. 9/20/2018

## Delta CO2

```{r plot-deltaCO2}
ggplot(data=pico1, aes(x = Days, y = deltaCO2, color = NamedStorm)) +
  geom_point() +
  labs(title = "DeltaCO2 vs. Days")
```

```{r calculations-max-DeltaCO2}
largest_deltaCO2 <- pico1 %>% 
  filter(!is.na(deltaCO2))%>% 
  select(deltaCO2, Date) %>% 
  arrange(desc(deltaCO2)) %>% 
  slice(1)
largest_deltaCO2
```

Maximum value happened right around Hurrricane Joaquin

## Bacteria

```{r plot-bacteriaMEAN}
ggplot(data=pico1, aes(x = Days, y = bacteriaMEAN, color = NamedStorm)) +
  geom_point() +
  labs(title = "BacteriaMEAN vs. Days")
```

```{r calculations-max-bacteria}
largest_bacteriaMEAN <- pico1 %>% 
  filter(!is.na(bacteriaMEAN))%>% 
  select(bacteriaMEAN, Date) %>% 
  arrange(desc(bacteriaMEAN)) %>% 
  slice(1)
largest_bacteriaMEAN
```

```{r plot-bacteria-vs-temperature}
p_temp <- ggplot(data=pico1, aes(x = TemperatureMEAN, y = bacteriaMEAN)) +
            geom_point() +
            labs(title = "BacteriaMEAN vs. TempMEAN") +
            geom_abline(intercept = 1340430.7, slope = 100283.6, color = "red", size = 1.5)
```

```{r plot-bacteria-vs-salinity-display}
p_salinity <- ggplot(data=pico1, aes(x = SalinityMEAN, y = bacteriaMEAN)) +
                geom_point() +
                labs(title = "BacteriaMEAN vs. SalinityMEAN")
p_temp+p_salinity
```

```{r linear-model-bacteria-temp}
bacteria_temp_model <- lm(bacteriaMEAN~TemperatureMEAN, data = pico1)
tidy(bacteria_temp_model)
```

```{r linear-model-bacteria-salinity}
bacteria_sal_model <- lm(bacteriaMEAN~SalinityMEAN, data = pico1)
tidy(bacteria_sal_model)
```

```{r correlation-bac-sal-temp}
temp_salinity <- pico1 %>% 
  select(bacteriaMEAN, SalinityMEAN, TemperatureMEAN)

ggcorr(temp_salinity, palette = "RdBu", label = TRUE)
```

The p-value for the linear regression shows that there is a relationship between mean bacteria and mean temperature, but there does not appear to be a strong correlation between the two

```{r plot-bacteria-vs-pH}
p_pH <- ggplot(data=pico1, aes(x = pHMEAN, y = bacteriaMEAN)) +
                geom_point() +
                labs(title = "BacteriaMEAN vs. pHMEAN") +
                geom_abline(intercept = 64645987, slope = -7658991, color = "red", size = 1.5)
```

```{r linear-model-bacteria-pH}
bacteria_ph_model <- lm(bacteriaMEAN~pHMEAN, data = pico1)
tidy(bacteria_ph_model)
```

```{r linear-model-bacteria-oxygen}
bacteria_oxygen_model <- lm(bacteriaMEAN~OxygenMEAN, data = pico1)
tidy(bacteria_oxygen_model)
```

```{r plot-bacteria-oxygen-display}
p_oxygen <- ggplot(data=pico1, aes(x=OxygenMEAN, y=bacteriaMEAN)) +
              geom_point() +
              labs(title = "BacteriaMEAN vs. OxygenMEAN") +
              geom_abline(slope = -541933, intercept = 7384236, color = "red", size = 1.5)
p_oxygen + p_pH
```
```{r correlation-bac-o-pH}
o_pH_bacteria <- pico1 %>% 
  select(bacteriaMEAN, OxygenMEAN, pHMEAN)

ggcorr(o_pH_bacteria, palette = "RdBu", label = TRUE)
```

Again, there does not appear to be a strong correlation. between mean bacteria. and pH and oxygen, despite the p-values of the linear models suggesting otherwise

\newpage

## Looking at pH alongside Oxygen

```{r display-ph-oxygen}
(p1 + scale_y_continuous(limits=c(7.25, 8.5)))+(p5 + scale_y_continuous(limits=c(3, 13)))
```

```{r plot-o-on-ph}
ggplot(pico1) +
  geom_line(aes(y = pHMEAN, x = Days),  size = 0.25) +
  geom_line(aes(y = OxygenMEAN, x = Days), color = "#c13d75ff")  +
  labs(title = "Mean Oxygen Plotted with Mean pH")
```

The mean oxygen measured by ProODO appears to follow a similar pattern of peaks as the mean pH calculated using CO2SYS.

```{r plot-ph-vs-oxygen}
ggplot(data = pico1, aes(x = OxygenMEAN, y = pHMEAN)) +
  geom_point() + 
  labs(title = "pHMEAN vs. OxygenMEAN")
```

```{r o-pH-model}
o_pH_model <- lm(pHMEAN ~ OxygenMEAN, data = pico1)
tidy(o_pH_model)
```



```{r correlation-o-pH}
o_pH <- pico1 %>% 
  select(OxygenMEAN, pHMEAN)

ggcorr(o_pH, palette = "RdBu", label = TRUE)
```


There appears to be a pretty strong positive correlation between the mean oxygen and pH levels.

```{r looking-at-temp}
p7 <- ggplot(data=pico1, aes(x=Days, y= TemperatureMEAN)) +
  geom_point() +
  labs(title = "TemperatureMEAN vs. Days")
p8 <- ggplot(data=pico1, aes(x=Days, y= TempSurfaceMEAN)) +
  geom_point() +
  labs(title = "TempSurfaceMEAN vs. Days")

p7+p8
```

```{r ggpairs-o-ph-temp}
o_pH_temp <- pico1 %>% 
  select(OxygenMEAN, pHMEAN, TemperatureMEAN)

ggpairs(o_pH_temp, columns = 1:ncol(o_pH_temp), title = "",  
  axisLabels = "show", columnLabels = colnames(o_pH_temp[]))
```

\newpage

# TIME SERIES

## pH Mean

```{r no-missing-ph-values}
full_pHMEAN <-  pico1 %>% 
  filter(!is.na(pHMEAN))
```


```{r time-series-pHMEAN}
pico1.ts <- ts(full_pHMEAN$pHMEAN, start = c(2010, 7), frequency = 52)
```

```{r decomposed-ph-ts}
pico1.Decomposed <- stl(pico1.ts, s.window = "periodic")
plot(pico1.Decomposed)
```

```{r changing-decomposed-dataframe}
pico1_Components <- as.data.frame(pico1.Decomposed$time.series[,1:3])
pico1_Components <- mutate(pico1_Components,
        Observed = full_pHMEAN$pHMEAN,     
        Days = full_pHMEAN$Days)
```

```{r plotting-decomposed-ts}
p_orig <- ggplot(pico1_Components) +
  geom_line(aes(y = Observed, x = Days),  size = 0.25) +
  geom_line(aes(y = trend, x = Days), color = "#c13d75ff") +
  ylab(expression("Mean pH calculated using CO2SYS")) + 
  scale_y_continuous(limits=c(7, 8.5)) +
  labs(title = "Trend Mapping onto Data")

p_seasonal <- ggplot(pico1_Components) +
  geom_line(aes(y = Observed, x = Days),  size = 0.25) +
  geom_line(aes(y = seasonal, x = Days), color = "#c13d75ff") +
  ylab(expression("Mean pH calculated using CO2SYS")) + 
  labs(title = "Seasonal Cycle Mappping onto Data")

p_orig+p_seasonal
```

```{r mann-kendall-ph}
SeasonalMannKendall(pico1.ts)
```

p-vale is. less. than 0.05, so we can reject the null hypothesis meaning that there is a trend

## Oxygen

```{r no-missing-o-values}
full_oMEAN <-  pico1 %>% 
  filter(!is.na(OxygenMEAN))
```


```{r time-series-oxygenMEAN}
pico1.ts_o <- ts(full_oMEAN$OxygenMEAN, start = c(2010, 7), frequency = 52)
```

```{r decomposed-oxygen}
pico1.Decomposed_o <- stl(pico1.ts_o, s.window = "periodic")
plot(pico1.Decomposed_o)
```

```{r oxygen-dataframe-edited}
pico1_Components_o <- as.data.frame(pico1.Decomposed_o$time.series[,1:3])
pico1_Components_o <- mutate(pico1_Components_o,
        Observed = full_oMEAN$OxygenMEAN,     
        Days = full_oMEAN$Days)
```

```{r plotting-decomposed-oxygen}
p_orig_o <- ggplot(pico1_Components_o) +
  geom_line(aes(y = Observed, x = Days),  size = 0.25) +
  geom_line(aes(y = trend, x = Days), color = "#c13d75ff") +
  ylab(expression("Mean Oxygen Measured by proODO")) + 
  labs(title = "Trend Mapping onto Data")

p_seasonal_o <- ggplot(pico1_Components_o) +
  geom_line(aes(y = Observed, x = Days),  size = 0.25) +
  geom_line(aes(y = seasonal, x = Days), color = "#c13d75ff") +
  ylab(expression("Mean Oxygen Measured by proODO")) + 
  labs(title = "Seasonal Cycle Mappping onto Data")

p_orig_o+p_seasonal_o
```

```{r mann-kendall-o}
SeasonalMannKendall(pico1.ts_o)
```
